---
title: "p8105_hw1_tb2715_commits"
author: "Tess"
date: "9/20/2019"
output: github_document
---

#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

#Problem 1

The chunk below creates a dataframe containing a sample of size 8 from a random normal distribution. Samples are classified as above or below 0 and the variable classification is explored between numeric and character variables.

```{r create dataframe}
eight_rand_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = norm_samp > 0,
  vec_char = c("I", "enjoy", "yoga", "cookies", "nature", "dogs", "sunshine", "family"),
  vec_factor = factor(c("orange", "orange", "orange", "orange","blue", "blue", "blue", "magenta" )))

mean(pull(eight_rand_df, norm_samp))

mean(pull(eight_rand_df, vec_logical))

mean(pull(eight_rand_df, vec_char))

mean(pull(eight_rand_df, vec_factor))
```

The mean of the samples cannot be computed for vec_char and vec_ factor. The argument is not numerical. 

```{r variable type manipulation, eval = FALSE}
eight_rand_df_V2 = tibble(
  norm_samp = rnorm(8),
  vec_logical = as.numeric(norm_samp > 0),
  vec_char = as.numeric(c("I", "enjoy", "yoga", "cookies", "nature", "dogs", "sunshine", "family")),
  vec_factor = as.numeric(factor(c("orange", "orange", "orange", "orange","blue", "blue", "blue", "magenta" ))))
```

Above, when the values are transformed to numerical values, the return is still NA as they cannot be transformed to a numerical value. NAs are introduced via coercion. 

```{r numeric variables vs factors}
numeric_eight_rand_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = as.numeric(norm_samp > 0), 
  norm_samp1 = (vec_logical * norm_samp)) 
  
 logical_eight_rand_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = factor(norm_samp > 0), 
  norm_samp2 = (vec_logical * norm_samp)) 

factor_and_numeric_eight_rand_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = factor(norm_samp > 0),
  vec_logical1 = as.numeric(vec_logical),
  norm_samp3 = (vec_logical1 * norm_samp)) 
```

Above, the classification of variables is manipulated to factors and numeric values. When a variable is classified as a factor, it cannot be used as a number in functions such as trying to multiply variables. When the logical value is a factor and then manipulated to be numeric, the values are different than when originally classified as numeric. 

#Problem 2
```{r create data frame for question 2}
plot_df = tibble(
  x = rnorm(500), 
  y = rnorm(500), 
  vec_logical = (x + y > 1), 
  num_vector = as.numeric(vec_logical), 
  factor_vector = factor(vec_logical))

#code for inline citations
mean_samp = mean(pull(plot_df, x))

median_samp = median(pull(plot_df, x))

stdev_samp = sd(pull(plot_df, x))

rows = nrow(plot_df)

columns = ncol(plot_df)

proportion = mean(pull(plot_df, vec_logical))
```

This vector contains `r round(rows)` and `r round(columns)`. Additionally, the mean of x is `r round(mean_samp, digits = 2)`, the median of x is `r round(median_samp, digits = 2)`, and the standard deviation of x is `r round(stdev_samp, digits = 2)`.

The proportion of cases for which x + y > 1 is `r round(proportion, digits = 2)`

```{r}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()

ggsave("plot_df.pdf", width = 4, height = 4)

ggplot(plot_df, aes(x = x, y = y, color = num_vector)) + geom_point()

ggplot(plot_df, aes(x = x, y = y, color = factor_vector)) + geom_point()
```

There is no difference in the points on the graph that are colored as above 1 or less than 1, but the scale is quite different. When the logical factor is classified as numeric, the scale extends from 0 to 1. Not relevant in this scenario where it is only classified as a 0 or 1, but it could demonstrate a range of values. When the logical vector is classified as a factor, the scale is determined by true or false and there are no additional options.

